---
title: "Portfolio 5"
---

I found an interesting looking dataset on the Tidy Tuesday site about age gaps in movies, and I want to use this portfolio to see if I can uncover some interesting patterns in it.

The link the dataset is here: https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-02-14/readme.md

Let's load what we need to start below:

```{r setup, message = FALSE}
library(tidyverse)

movies <- read.csv("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p05/movies.csv")
```

The first question that jumps out: what was the movie with the greatest age gap between actors playing love interests?

```{r greatest-gap}
max_diff <- movies$Age.Difference %>%
  max()
max_diff

movies %>%
  filter(Age.Difference == max_diff) %>%
  select(Movie.Name, Release.Year)
```
Similarly, we can see which movies have no age gap between leads:

```{r no-gap}
movies %>%
  filter(Age.Difference == 0) %>%
  select(Movie.Name, Release.Year)
```

Another question: how have age gaps in movies changed over time? I can imagine that really huge age gaps might have been more common in the past than they are today, and it would be interesting to visualize that change over time:

```{r gap-over-time}
movies %>%
  ggplot(aes(x = Release.Year, y = Age.Difference)) + 
  geom_jitter() +
  geom_smooth(method = "lm")
```








