---
title: "Portfolio 9"
---
> I want to use this portfolio to play around with some methods of visualizing the thesis data I've collected so far. I've already visualized some of it in previous projects, but I figure it will come in handy later to have taken some dedicated time to think about how to present my data graphically


> Let's get everything loaded and prepped:

```{r setup}
library(tidyverse)
library(readxl)
```

```{r load-data}
day_one <- read_excel("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p03/IC Diet Choice.xlsx",
                      sheet = 10,
                      col_names = TRUE)
day_two <- read_excel("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p03/IC Diet Choice.xlsx",
                      sheet = 11,
                      col_names = TRUE)
day_three <- read_excel("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p03/IC Diet Choice.xlsx",
                      sheet = 12,
                      col_names = TRUE)
day_four <- read_excel("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p03/IC Diet Choice.xlsx",
                      sheet = 13,
                      col_names = TRUE)
day_five <- read_excel("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p03/IC Diet Choice.xlsx",
                      sheet = 14,
                      col_names = TRUE)
day_six <- read_excel("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p03/IC Diet Choice.xlsx",
                      sheet = 15,
                      col_names = TRUE)
day_seven <- read_excel("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p03/IC Diet Choice.xlsx",
                      sheet = 16,
                      col_names = TRUE)
avg_kcal <- day_seven <- read_excel("/Users/lawilson1999/Desktop/Data Science Spring 2023/Portfolio/p03/IC Diet Choice.xlsx",
                      sheet = 9,
                      col_names = TRUE)

columns <- c('rat', 'zero', 'twenty', 'forty', 'sixty', 'eighty', 'hundred', 'water')
colnames(day_one) <- columns
colnames(day_two) <- columns
colnames(day_three) <- columns
colnames(day_four) <- columns
colnames(day_five) <- columns
colnames(day_six) <- columns
colnames(day_seven) <- columns
colnames(avg_kcal) <- columns
```

```{r concatenation-and-kcal-conversion}
day_one <- day_one %>%
  mutate(day = 1)
day_two <- day_two %>%
  mutate(day = 2)
day_three <- day_three %>%
  mutate(day = 3)
day_four <- day_four %>%
  mutate(day = 4)
day_five <- day_five %>%
  mutate(day = 5)
day_six <- day_six %>%
  mutate(day = 6)
day_seven <- day_seven %>%
  mutate(day = 7)

days <- list(day_one, day_two, day_three, day_four, day_five, day_six, day_seven)

all_days <- do.call("rbind", days)

all_days_kcal <- all_days
all_days_kcal$zero <- all_days_kcal$zero*9
all_days_kcal$twenty <- all_days_kcal$twenty*8
all_days_kcal$forty <- all_days_kcal$forty*7
all_days_kcal$sixty <- all_days_kcal$sixty*6
all_days_kcal$eighty <- all_days_kcal$eighty*5
all_days_kcal$hundred <- all_days_kcal$hundred*4
```

> And with that out of the way we can get started!

## Change Over Time

> One of the arguments I'm making for the length of time we chose to train the animals is that it would hopefully result in stable food preferences by the time we got to testing.  Let's see if that's true:

```{r change-over-time}
# Reformat data frames for ggplot
all_days_kcal_long <- all_days_kcal %>%
  pivot_longer(cols = -c(rat, day),
               names_to = "diet",
               values_to = "kcals")

avg_kcal_long <- avg_kcal %>%
  pivot_longer(cols = -c(rat),
               names_to = "diet",
               values_to = "kcals")

# Average intake over time
all_days_kcal_long %>%
  filter(rat == "Avg") %>%
  filter(diet != "water") %>%
  ggplot(aes(x = day, y = kcals, color = diet)) +
  geom_line()
    
```