---
title: "Portfolio 10"
---

> A few months ago, we spent some time in class talking about numbats, a very interesting looking marsupial species that is native to Australia and currently considered to be endangered. Today, I'm going to answer some questions about them by analyzing a dataset containing information on numbbat sightings.

> The Tidy Tuesday data I'll be working with is linked here: https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-03-07/readme.md

```{r setup, message=FALSE}
library(tidyverse)

tuesdata <- tidytuesdayR::tt_load(2023, week = 10)
numbats <- tuesdata$numbats
```

> The data actually comes with some questions that I'm going to try to answer. No guarantee I'll get all of them, but it'll be fun to try.

## Where do you find numbats in Australia?

> The data include longitude and latitude info, so we can just plot those against each other to see where numbats have been sighted in Australia:

```{r location}
numbats %>%
  ggplot(aes(x = decimalLongitude, y = decimalLatitude)) +
  geom_point(color = "orange")
```

> This by itself isn't incredibly intuitive, but I can't seem to get the points to cleanly superimpose over a map of Australia. I've checked though, and the sightings do span the entirety of the country.

## What time of day do numbat sightings occur?

> A density plot of the "hour" variable should tell us!

```{r hour}
numbats %>%
  ggplot(aes(x = hour)) +
  geom_density(fill = "orange")

numbats %>%
  drop_na(hour)%>%
  summarize(mean(hour))
```

> Looks like they come out just after noon!

## Are numbat sightings more frequent in the summer or winter?

> In Australia, summer covers the months of December, January, and February, while winter ranges from June to August. Lets use that knowledge to slice and dice this dataset the way we want:

```{r numbat-seasons}
#Processing
numbats_szns <- numbats %>%
  filter(month == c("Dec","Jan", "Feb", "Jun", "Jul", "Aug")) %>%
  mutate(season = case_when(month == "Dec" ~ "summer",
                            month == "Jan" ~ "summer",
                            month == "Feb" ~ "summer",
                            month == "Jun" ~ "winter",
                            month == "Jul" ~ "winter",
                            month == "Aug" ~ "winter",))

numbat_szns_counts <- numbats_szns %>%
  group_by(season) %>%
  summarize("count" = n())

# Visualization
numbats_szns %>%
  ggplot(aes(x = season, fill = season)) + 
  geom_histogram(stat = "count", alpha = 0.7)
```

> Looks like numbat sightings are pretty substantially more common in the summer!

## Do sightings happen more on week days than weekends?

> This should also be pretty straightforward as well:

```{r week}
numbats %>%
  drop_na(wday) %>%
  ggplot(aes(x = factor(wday, level = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")), fill = wday)) + 
  geom_histogram(stat = "count", alpha = 0.7)
```

> Seems like they like weekdays, and particularly Mondays!